---
title: "Lab 02 - Plastic waste"
author: "Noah Booker"
date: "01/29/25"
output: github_document
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse)
```

```{r load-data, message=FALSE}
plastic_waste <- read.csv("data/plastic-waste.csv")
```

## Exercises

### Exercise 1

Below are histograms plotting the distribution of plastic waste per capita faceted by continent.
The majority of all continents' distributions of plastic waste per capita fall between 0 and .5 kg/day. For those continents which have countries with a plastic waste per capita above .5 kg/day, it seems that North America has the most followed by South America and Asia tied for second place.

```{r plastic-waste-continent}
ggplot(data = plastic_waste, aes(x = plastic_waste_per_cap)) +
    geom_histogram(binwidth = 0.2) + facet_wrap(~continent)
```

### Exercise 2

Below, we define the color and fill of the curves by mapping aesthetics of the plot but define the alpha level as a characteristic of the plotting geom because we want color and fill to vary according to the data (by continent) but want the alpha level (transparency) to be applied the same way all the density curves.
```{r plastic-waste-density}
ggplot(
  data = plastic_waste,
  mapping = aes(
    x = plastic_waste_per_cap,
    color = continent,
    fill = continent
  )
) +
  geom_density(alpha = 0.2)
```

### Exercise 3

Remove this text, and add your answer for Exercise 3 here.

```{r plastic-waste-violin}
# insert code here
```

### Exercise 4

Remove this text, and add your answer for Exercise 4 here.

```{r plastic-waste-mismanaged}
# insert code here
```

```{r plastic-waste-mismanaged-continent}
# insert code here
```

```{r plastic-waste-population-total}
# insert code here
```

```{r plastic-waste-population-coastal}
# insert code here
```

### Exercise 5

Remove this text, and add your answer for Exercise 5 here.

```{r recreate-viz}
# insert code here
```


