---
title: "Lab 02 - Plastic waste"
author: "Noah Booker"
date: "01/29/25"
output: github_document
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse)
```

```{r load-data, message=FALSE}
plastic_waste <- read.csv("data/plastic-waste.csv")
```

## Exercises

### Exercise 1

Below are histograms plotting the distribution of plastic waste per capita faceted by continent.
The majority of all continents' distributions of plastic waste per capita fall between 0 and .5 kg/day. For those continents which have countries with a plastic waste per capita above .5 kg/day, it seems that North America has the most followed by South America and Asia tied for second place.

```{r plastic-waste-continent}
ggplot(data = plastic_waste, aes(x = plastic_waste_per_cap)) +
    geom_histogram(binwidth = 0.2) + facet_wrap(~continent)
```

### Exercise 2

Below, we define the color and fill of the curves by mapping aesthetics of the plot but define the alpha level as a characteristic of the plotting geom because we want color and fill to vary according to the data (by continent) but want the alpha level (transparency) to be applied the same way all the density curves.
```{r plastic-waste-density}
ggplot(
  data = plastic_waste,
  mapping = aes(
    x = plastic_waste_per_cap,
    color = continent,
    fill = continent
  )
) +
  geom_density(alpha = 0.2)
```

### Exercise 3

Unlike the box plots, the violin plots show the shape of the distribution. This allows one to visualize, for example, whether the distribution is multimodal (more than one peak).
Unlike a violin plot, box plots explicitly mark the first and third quartiles, the median, and outliers of each distribution.

```{r plastic-waste-violin}
ggplot(
  data = plastic_waste,
  mapping = aes(
    x = continent,
    y = plastic_waste_per_cap
  )
) +
  geom_violin()
```

### Exercise 4

The scatterplot below visualizes the relationship between plastic waste per capita and mismanaged plastic waste per capita. There seems to be an overall positive correlation. Also, there seems to be more variance in mismanaged plastic waste than in overall plastic waste (though they are visualized in different scales, so this may be misleading).

```{r plastic-waste-mismanaged}
ggplot(data = plastic_waste, mapping = aes(x = plastic_waste_per_cap, y = mismanaged_plastic_waste_per_cap)) + geom_point()
```

There seems to be an overall postive correlation between plastic waste per capita and mismanage plastic waste per capita. However, some continents seem to have a larger correlation. For instance, when comparing Africa to North America, it seems that as plastic waste per capita increases, mismanaged plastic saste per capita increases more quickly in Africa than it does in North America.
```{r plastic-waste-mismanaged-continent}
ggplot(data = plastic_waste, mapping = aes(x = plastic_waste_per_cap, y = mismanaged_plastic_waste_per_cap, color = continent)) + geom_point()
```

There seems to be more variance in countries' coastal populations than there is countries' total population, but there doesn't seem to be any notable differences in the relationship between the two of these variables and plastic waste per capita.
```{r plastic-waste-population-total}
ggplot(data = plastic_waste, mapping = aes(x = plastic_waste_per_cap, y = total_pop, color = continent)) + geom_point()
```

```{r plastic-waste-population-coastal}
ggplot(data = plastic_waste, mapping = aes(x = plastic_waste_per_cap, y = coastal_pop, color = continent)) + geom_point()
```

### Exercise 5

Remove this text, and add your answer for Exercise 5 here.

```{r recreate-viz}
# insert code here
```


